---
title: "TA543"
output: html_document
---

```{r}
dataset = read.csv('https://raw.githubusercontent.com/tomasdgus/finalProject2020/master/dataset543.csv')
```

```{r}
dataset$RegionNumber <- as.numeric(dataset$RegionNumber)
str(dataset)
```


```{r}
library(magrittr)
sort(unique(dataset$STATE))%>%head(20)
```

```{r, eval=TRUE}
USmap="https://raw.githubusercontent.com/tomasdgus/finalProject2020/master/states.json"
```

```{r, eval=TRUE}
library(geojsonio)
PROJmap="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
USzipMap=topojson_read(USmap,crs=PROJmap,stringsAsFactors = FALSE)
```

```{r, eval=TRUE}
object.size(USzipMap)
```
```{r}
library(rmapshaper)
USzipMap=ms_simplify(USzipMap)
```


```{r, eval=TRUE}
object.size(USzipMap)
```


```{r, eval=TRUE}
library(ggplot2)
ggplot(data=USzipMap) + geom_sf()
```


```{r}
names(USzipMap)
```


```{r}
str(USzipMap$STATE_NAME)
```


```{r}
layerUSMap=merge(USzipMap,dataset, 
                   by.x='STATE_NAME', 
                   by.y='STATE',
                   all.x=F) # if no coincidence don't keep shape.
```


```{r, eval=TRUE,warning=FALSE}
# This will make just a border of the state
baseMap <- ms_dissolve(USzipMap)
```


This is the base map, which may help us show the missing shapes.
```{r, eval=TRUE}
ggplot(baseMap) + geom_sf()
```


```{r, eval=TRUE,warning=FALSE,message=FALSE,results='hide'}
library(RColorBrewer)
library(tmap)
```


2. Decide the amount of groups:


```{r, eval=TRUE}
numberOfClasses = 5
```


3. Decide colors for each group of shapes (I chose a palette from [here](http://colorbrewer2.org/)):
```{r, eval=TRUE}
colorForScale='Greens' # color scale
RegionColor = 'Pastel1'
```



Map of the Region Breakdown
```{r, eval=TRUE}
layerUSMap1 = layerUSMap
layerUSMap2 = layerUSMap
layerUSMap2$cut=cut_number(layerUSMap$Confidence,4,
                            ordered_result=T,
                            dig.lab=4)
layerUSMap3 = layerUSMap
layerUSMap4 = layerUSMap
layerUSMap5 = layerUSMap
```

```{r}
baseLayer1 = ggplot(data=baseMap) + geom_sf()
layer1 = baseLayer1 +geom_sf(data = layerUSMap1, aes(fill=REGION),show.legend = T)+
    scale_fill_brewer(palette = RegionColor,name="Regions")
layer1 = layer1 + theme_classic() + theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())+labs(title.text = "Regions of the U.S.",caption.text = "Source: U.S. National Park Service")+theme(plot.title = element_text(hjust=0.75),plot.caption = element_text(hjust=1.5))
layer1
```


```{r}
#dataset$STATE <- dataset$ï..STATE
#dataset$STATE <- as.character(dataset$ï..STATE)
str(dataset)
dataset$GDPperCapita <- as.numeric(dataset$GDPperCapita)
dataset$InsuredPCT <- as.numeric(1-dataset$UninsuredPCT)
dataset$HSDropOut <- as.numeric(1-dataset$HighSchoolGradPct)
```


```{r}
#sorting region by violent crimer per 1000
library(magrittr)
library(ggplot2)
dataset[order(-dataset$ViolentCrimePer1000),c('REGION','ViolentCrimePer1000')]%>%head(20)
```

High School Graduation Rates By Region
```{r}
library(forcats)
basehsgr = ggplot(data = dataset, aes(x = fct_reorder(REGION, HighSchoolGradPct),y = HighSchoolGradPct),
                  xlab = ("Region"))
basehsgr = basehsgr + geom_boxplot() + labs(title = "High School Graduation Rates By Region")
basehsgr = basehsgr + theme_classic() + theme(axis.title.x = element_blank(), plot.title = element_text(hjust = 0.5, face = "bold"))  
basehsgr
```

Uninsured Rates By Region
```{r}
baseinsurex = ggplot(data = dataset,
                    aes(x = fct_reorder(REGION, Uninsured), y = Uninsured))
boxinsurex = baseinsurex + geom_boxplot()
boxinsurex = boxinsurex + theme_classic() + theme(axis.text.x = element_text(angle = 45,
                                                   hjust = 1,
                                                   size = 8),axis.title.x = element_blank()) + labs(title = "Uninsured Rates By Region", caption = "Source: US Census Bureau Current Population Survey") + theme(plot.title = element_text(hjust = 0.5, face = "bold"))
boxinsurex
```

Main Measures By Region
```{r}
basegdp = ggplot(data = dataset[dataset$GDPperCapita <=120000,],
              aes(x= fct_reorder(REGION, GDPperCapita), y = GDPperCapita))
boxgdp = basegdp + geom_boxplot()
boxgdp = boxgdp + theme_classic() + theme(axis.text.x = element_text(angle = 45,
                                                   hjust = 1,
                                                   size = 8),axis.title.x = element_blank()) + labs(caption = "Source: Bureau of Economic Analysis")
basecrime = ggplot(data = dataset[dataset$ViolentCrimePer1000<=10,],
                   aes(x=fct_reorder(REGION, ViolentCrimePer1000), y = ViolentCrimePer1000))
boxcrime = basecrime + geom_boxplot()
boxcrime = boxcrime + theme_classic() + theme(axis.text.x = element_text(angle = 45,
                                                   hjust = 1,
                                                   size = 8),axis.title.x = element_blank()) + labs(caption = "Source: FBI State Crime Statistics")
baseinsure = ggplot(data = dataset,
                    aes(x = fct_reorder(REGION, Uninsured), y = Uninsured))
boxinsure = baseinsure + geom_boxplot()
boxinsure = boxinsure + theme_classic() + theme(axis.text.x = element_text(angle = 45,
                                                   hjust = 1,
                                                   size = 8),axis.title.x = element_blank()) + labs(caption = "Source: US Census Bureau Current Population Survey")
baseeduc = ggplot(data = dataset,
                  aes(x = fct_reorder(REGION, BachelorPct), y = BachelorPct))
boxedu = baseeduc + geom_boxplot()
boxedu = boxedu + theme_classic() + theme(axis.text.x = element_text(angle = 45,
                                                   hjust = 1,
                                                   size = 8),axis.title.x = element_blank()) + labs(caption = "Source: US Census Bureau")
annotate_figure(ggarrange(boxgdp, boxcrime, boxinsure, boxedu),top = text_grob("Main Measures by Region\n",size= 14,hjust=0.5, face = "bold"))
```

What Is Related to Crime?
```{r}
#plot show the correlation between Violent Crime and Uninsured Rates
library(tidyverse)
baseregioncrime = ggplot(dataset[dataset$ViolentCrimePer1000<=10,], aes(x=UninsuredPCT, y=ViolentCrimePer1000))
plotcrime = baseregioncrime + geom_point()+theme_classic() + labs(caption = "Source: US Census Bureau Current Population Survey")+ geom_smooth(method="lm",se=FALSE)


baseregioncrime2 = ggplot(dataset[dataset$ViolentCrimePer1000<=10&dataset$GDPperCapita<=120000,], aes(x=GDPperCapita, y=ViolentCrimePer1000))
plotgdpcrime = baseregioncrime2 + geom_point()+theme_classic() + labs(caption = "Source: US Bureau of Economic Analysis")+ geom_smooth(method="lm",se=FALSE)



baseregioncrime3 = ggplot(dataset[dataset$ViolentCrimePer1000<=10&dataset$HighSchoolGradPct>=.80,], aes(x=HighSchoolGradPct, y=ViolentCrimePer1000))
plothscrime = baseregioncrime3 + geom_point()+theme_classic() + labs(caption = "Source:US Census Bureau")+ geom_smooth(method="lm",se=FALSE)


crimeplots = annotate_figure(ggarrange(plotcrime,plotgdpcrime,plothscrime), top ="",)
crimeplots + labs(title = "What Is Related to Crime?") + theme(plot.title = element_text(face = "bold"))
```

A Closer Look at Crime with Uninsured Percentages
```{r}
baseregioncrime = ggplot(dataset[dataset$ViolentCrimePer1000<=10,], aes(x=UninsuredPCT, y=ViolentCrimePer1000))
plotcrime = baseregioncrime + geom_point()+theme_classic() + labs(title = "Is No Healthcare Related to Crime?",caption = "(Excluding Washington DC) Source: US Census Bureau Current Population Survey") + geom_smooth(method="lm",se=FALSE)
plotcrime = plotcrime + theme(plot.title = element_text(hjust = 0.5, face = "bold"))
plotcrime = plotcrime + annotate("text",x= 0.18,y=6.4,label="Alaska")+
                        annotate("text",x=0.06,y=1,label="Vermont")
plotcrime
```

State Confidence
```{r}
pacificx = c(-110,-125)
pacificy = c(31,50)
intermountainx =c(-115,-94)
intermountainy = c(27,50)
midwestx = c(-104,-81)
midwesty = c(33,50)
southeastx = c(-94,-75)
southeasty = c(20,38)
northeastx = c(-80,-68)
northeasty = c(40,48)
```


```{r}
baseLayer2=ggplot(data = baseMap) +geom_sf() 
layer2 = baseLayer2 + geom_sf(data = layerUSMap2, aes(fill=cut),show.legend = T) +
                 scale_fill_brewer(palette = colorForScale,
                                   name = "Confidence")+theme_classic()
layer2 = layer2 + theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())+labs(title.text="State Confidence")+theme(plot.title = element_text(hjust=0.75))
layer2 
```

```{r}
layerUSMap3$cut=cut_number(layerUSMap$Uninsured,2,
                            ordered_result=T,
                            dig.lab=2)
baseLayer3=ggplot(data = baseMap) +geom_sf() 
layer3 = baseLayer3 + geom_sf(data = layerUSMap3, aes(fill=cut),show.legend = T) +
                 scale_fill_brewer(palette = 'PuBu',
                                   name = "Uninsured Rate")+theme_classic()
layer3 = layer3 + theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())+labs(title = "Uninsured Rate",caption= "Source:US Census Bureau")+theme(plot.title = element_text(hjust=0.75))
layer3 
```

```{r}
layerUSMap4$cut=cut_number(layerUSMap$HSDropOut,2,
                            ordered_result=T,
                            dig.lab=2)
baseLayer4=ggplot(data = baseMap) +geom_sf() 
layer4 = baseLayer4 + geom_sf(data = layerUSMap4, aes(fill=cut),show.legend = T) +
                 scale_fill_brewer(palette = 'PuBu',
                                   name = "HS Drop Out %")+theme_classic()
layer4 = layer4 + theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())+labs(title = "High School Drop Out Percentage",caption= "Source:US Census Bureau")+theme(plot.title = element_text(hjust=0.75))
layer4 
```
```{r}
layerUSMap5$cut=cut_number(layerUSMap$ViolentCrimePer1000,2,
                            ordered_result=T,
                            dig.lab=4)
baseLayer5=ggplot(data = baseMap) +geom_sf() 
layer5 = baseLayer5 + geom_sf(data = layerUSMap5, aes(fill=cut),show.legend = T) +
                 scale_fill_brewer(palette = 'PuBu',
                                   name = "Violent Crime per Capita")+theme_classic()
layer5 = layer5 + theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())+labs(title = "Violent Crime per Capita",caption= "Source:US Census Bureau")+theme(plot.title = element_text(hjust=0.75))
layer5 
```
```{r}
ggarrange(layer3,layer4,layer5)
```


```{r}
pacific2 = layer2 + coord_sf(xlim = pacificx ,ylim=pacificy)+theme(legend.position = 'none')
pacific2
intermountain2 = layer2 + coord_sf(xlim=intermountainx, ylim=intermountainy)
intermountain2
midwest2 = layer2+coord_sf(xlim=midwestx,ylim=midwesty)
midwest2
northeast2 = layer2+coord_sf(xlim=northeastx,ylim=northeasty)
northeast2
```
```{r}
Confidencehistobase = ggplot(data=dataset, aes(x=fct_reorder(REGION,Confidence), y=Confidence))
Confidencehisto = Confidencehistobase + geom_boxplot()+theme_classic() + theme(axis.title.x = element_blank())
Confidencehisto
```

```{r}
library(ggpubr)
mixed2 = ggarrange(pacific2,northeast2,ncol=2,nrow=1)
mixed2full = annotate_figure(mixed2,top = "Pacific vs. Northeast Region")
mixed2full
```
